### 2020-09-20 manuscript/paper_BRM.Rmd
add some references and remove two-state sim

library(papaja)
library(tidyverse)
library(signal)
library(CircStats)
library(depmixS4)
library(psych)
library(here)
source(here("algorithm/preprocessing_helper_functions.R"))
source(here("algorithm/model_helper_functions.R"))

@@

On one hand, many of these classification algorithms use non-parametric methods to differentiate between eye movement events. For instance, the classic "Velocity-Threshold" algorithm classifies samples with a velocity above a fixed threshold as saccades [@Stample1993]. On the other hand, many parametric methods have been developed recently. Some of them require human labeled training data as input and can therefore be termed as supervised. For example, @Bellet2019 trained a convolutional neural network (CNN) on eye-tracking data from humans and macaques and achieved classification results that were similarly accurate to those of human coders. However, these supervised methods have several disadvantages: First, they require labeled data (usually labeled by humans) as input. Second, they often rely on hyperparameters which are difficult to theoretically justify. Third, their parameters are usually difficult to interpret: Most supervised methods operate in a black-box fashion, which implies that they do not make any explicit assumptions about the data generating process. Beyond that, there are several algorithms using unsupervised techniques. They typically do not need to be trained on labeled data and are also more flexible in classifying data from different individuals, tasks, or eye-trackers.  
Besides discriminating between supervised and unsupervised methods, algorithms can vary in whether they are explicitly modeling the data generating process and are thus able to simulate new data. To our knowledge, these *generative models* have been rarely used to classify eye movement data [cf. @Mihali2017, @Wadehn2019]. However, they come with various advantages and we will summarize the two most important ones for eye movement classification here [for an overview, see @Epstein2008].  
First, generative models allow to not only predict (classify) but also explain eye-tracking phenomena. For instance, they can inhere parameters that explain why oscillations only occur after but not before saccades. When the goal is to understand eye movements and improve their classification based on this understanding, this aspect is clearly an advantage over non-parametric or supervised methods. Second, generative models can challenge prevailing theoretical assumptions and bring up new research questions [@Epstein2008]. For example, they might suggest that oscillations also occur before saccadic eye movements [as has been mentioned by @Nystrom2010] or that the assumption of discrete eye movements does not hold [as has been discussed in @Andersson2017].  

@@

Several studies have tested the performance of HMM algorithms against other classification methods: I-HMM has been deemed as robust against noise, behaviorally accurate, and showing a high sample-to-sample agreement to human coders [@Andersson2017; @Komogortsev2010; @Salvucci2000]. However, the agreement was lower when compared to an algorithm using a Bayesian mixture model [@Kasneci2014; @Tafaj2012]. NSLR-HMM showed even higher agreement to human coding than I-HMM but was outperformed by the CNN algorithm by @Bellet2019 [@Pekkanen2017].  

@@

[^1]: @Hessels2017 called step two the *search rule* and step three the *classifcation rule*. For non-parametric methods, this distinction might be accurate. However, for parametric methods, we think calling step two *classification* is more appropriate since the probabilistic classification is done here. Step three usually consist of some heuristic relabeling and correcting for classification errors.

## Preprocessing
Algorithms require variables that describe gaze data (hereafter called *eye movement metrics*) to classify them into events. Many eye movement metrics have been proposed and used in previous algorithms [for examples, see @Zemblys2018; @Andersson2017]. However, many of them rely on thresholds or window ranges that have to be set by the user [e.g., the distance between the mean position in a 100 ms window before and after each sample, see @Olsson2007]. This can be problematic because such parameters are often set without theoretical justification and differ substantially between metrics or heavily depend on the eye-tracker's sampling frequency [@Andersson2017]. In gazeHMM, we used velocity, acceleration, and sample-to-sample angle [synonymous to relative or change in angle; @Larsson2013] because they belong to the most basic metrics which do not require parameter settings.  
Theoretically, these three metrics should separate eye movement events clearly. Fixations typically inherit samples with low velocity and acceleration [@Larsson2013]. Due to tremor, we assume that the angle between samples should not follow any direction but a random walk [@Duchowski2017; @Larsson2015]. In contrast, saccade samples usually have a high velocity and acceleration and roughly follow the same direction. PSO samples tend to have moderate velocity and high acceleration since they occur between saccades and low velocity events [@Larsson2013]. They can be specifically distinguished by their change in direction clustered around 180 degrees [@Pekkanen2017]. Importantly, the oscillations depend on the resolution of the gaze recording: Eye-trackers with higher sampling frequency yield more changes in direction and more samples in between those changes. Those samples in between typically follow the same direction. Thus, with high sampling frequencies, PSO samples might also cluster around a sample-to-sample angle of zero with outliers around 180 degrees. Lastly, smooth pursuit samples have a moderate velocity but low acceleration (due to the smoothness) and like saccades they follow the same direction [@Larsson2013; @Leigh2015]. Other algorithms focus exclusively on classifying microsaccades [e.g., @Mihali2017], but as stated earlier, these events were not in the scope of gazeHMM.  
The velocity and acceleration signals are computed from the raw gaze position by using a Savitzky-Golay filter [@Savitzky1964; similar to @Nystrom2010]. The sample-to-sample angle is calculated as: $$\alpha(t) = \arctan\left(\frac{y_{t+1} - y_{t}}{x_{t+1} - x_{t}}\right) -\arctan\left(\frac{y_{t} - y_{t-1}}{x_{t} - x_{t-1}}\right),$$ with $\alpha(t) := \alpha(t) + 2\pi$ for $\alpha(t) < 0$, and is therefore bound between 0 and $2\pi$.  
Most of the missing data in eye movement classification are due to blinks. Therefore, the user can indicate which samples should be labeled as blinks (other missing samples are treated as noise). Often, eye-trackers record a few samples with unreasonably high velocity and acceleration before losing the pupil signal when a blink occurs. Since these samples could distort the classification of saccades in the HMM, gazeHMM removes them heuristically. Before classifying the samples, it sets all samples within 50 ms before and after blink samples as missing. The window of 50 ms was rather motivated empirically than theoretically and can bet set to any value the user considers appropriate.

@@

When we developed gazeHMM, the gamma distribution appeared to fit eye movement data well, but we also note that it might not necessarily be the best fitting distribution for every type of eye movement data. We assume that the best fitting distribution will depend on the task, eye-tracker, and individual (see also the discussion).  

@@

Both the distributions and the metric operate on the full unit circle (i.e., between 0 and $2\pi$), which should lead to rather symmetric distributions. Because the fixations change their direction similar to a random walk, their sample-to-sample angle can be modeled by a uniform distribution. Thus, the uniform distribution should distinguish fixations from the other events.  

@@

## Postprocessing
After classifying gaze samples into states, gazeHMM applies a postprocessing routine to the estimated state sequence. We implemented this routine because constraining the transition probabilities for PSOs to turn into non-saccade events to zero often caused PSOs not to appear in the state sequence at all. Moreover, gazeHMM does not explicitly control the duration of events in the HMM which occasionally led to unreasonably short events. Thus, the postprocessing routine heuristically compensates for such violations. This routine relabels one-sample fixations and smooth pursuits, saccades with a duration below a minimum threshold, and PSOs that follow non-saccade events. Samples are relabeled as the state of the previous event. Finally, samples initially indicated as missing are labeled as noise (including blinks) and event descriptives are computed (e.g., fixation duration).  
The algorithm is implemented in R [version: 3.6.3; @RCoreTeam2020] and uses the packages \texttt{signal} [@Ligges2015] to compute velocity and acceleration signals, \texttt{depmixS4} [@Visser2010] for the HMM, \texttt{CircStats} [@Lund2018] for the von Mises distribution, and \texttt{BB} [@Varadhan2009] for Barzilai-Borwein spectral projected gradient optimization. The algorithm is available on GitHub (www.github.com/maltelueken/gazeHMM).

@@

## Results
### Parameter Variation
In the first part of the simulation, we examined how varying the parameters in the HMM affected the deviation of estimated parameters and accuracy of estimated state sequences. Figure \@ref(fig:plot-rmdspd-1-2) displays the RMdSPD between true and estimated parameters depending on which parameter has been manipulated in the HMM. The RMdSPD was below 0.1 for all estimated and manipulated parameters, indicating good recovery [^2]. The regression lines between manipulated true and estimated parameters are shown in Figures \@ref(fig:plot-regtr-1-2) and \@ref(fig:plot-regresp-1-2). With one outlier at parameter $\alpha_{acc;1}$, the estimated parameters matched the true parameters very closely. Thus, parameter change was captured well and the estimation almost unbiased. Considering accuracy, Figure \@ref(fig:plot-acc-1-2) displays Cohen's kappa between true and estimated hidden state sequences. With two exceptions, kappa values were almost one, suggesting nearly perfect agreement.

[^2]: Note that the initial state probability $\rho_i$ has $\textrm{RMdSPD} = 1$. Since the HMM only simulated one state sequence, this parameter is always either zero or one (leading to $\textrm{RMdSPD} = 1$). Therefore, we decided not to include it in the analysis.

(ref:plot-rmdspd-1-2) RMdSPD between true and estimated parameters of the two-state HMM in part one of the simulation. Labels on the x-axis indicate which true parameters have been manipulated and labels on the y-axis show for which estimated parameter the RMdSPD is displayed. Top facet labels specify in which state the parameters have been varied and right facet labels denote to which state estimated parameters belong. $\rho_i$ is the initial probability for state *i* (indicated by the right facet label), $a_{ij}$ is the probability to transition from state *i* to state *j*, $\alpha$ and $\beta$ are the shape and scale parameters of the gamma distributions, and $\mu$ and $\kappa$ are the mean and concentration parameter of the von Mises distribution.

```{r plot-rmdspd-1-2, fig.cap="(ref:plot-rmdspd-1-2)"}

print(plots.rmsd.1[[1]])

(ref:plot-regtr-1-2) Regression lines between true and estimated transition probabilities for the two-state HMM in part one. Right facet labels show *from* and top facet labels show *to* which state the HMM is moving. Dashed lines refer to perfect recovery.

```{r plot-regtr-1-2, fig.cap="(ref:plot-regtr-1-2)"}

print(plots.lm.tr[[1]])

(ref:plot-regresp-1-2) Regression lines between true and estimated response parameters of the two-state HMM in part one. Top facet labels indicate response parameters. Dashed lines refer to perfect recovery.

```{r plot-regresp-1-2, fig.cap="(ref:plot-regresp-1-2)"}

print(plots.lm.resp[[1]] + theme(axis.text = element_text(size = 10)))

(ref:plot-acc-1-2) Boxplots displaying Cohen's kappa depending on which parameter of the two-state HMM has been manipulated in part one. Top facet labels indicate for which state parameters have been manipulated. Black solid lines symbolize medians. Crosses represent outliers (distance to first/third quartile higher than 1.5 times the inter-quartile range [IQR]).

```{r plot-acc-1-2, fig.cap="(ref:plot-acc-1-2)"}

print(plots.acc.1[[1]])

For the HMM with three states, the RMdSPD is shown in Figure \@ref(fig:plot-rmdspd-1-3). When response parameters (other than $a_{i=j}$) were manipulated, the RMdSPDs for $a_{12}$ and $a_{31}$ were consistently between 0.1 and 0.5. Varying $\kappa$ in states two and three led to RMdSPDs between 0.1 and 0.5 in the respective states, which we interpreted as moderate recovery. Otherwise, RMdSPDs were consistently lower than 0.1, indicating good recovery. Inspecting the regression lines between true and estimated parameters (see Figures \@ref(fig:plot-regtr-1-3) and \@ref(fig:plot-regresp-1-3)) revealed strong and unbiased linear relationships (intercepts close to zero and slopes close to one). In contrast to the two-state HMM, larger deviations and more outliers were observed. Cohen's kappa values are presented in Figure \@ref(fig:plot-acc-1-3). For most estimated models, the kappa values between true and estimated state sequences were above 0.95, meaning almost perfect agreement. However, for some models, we observed kappas clustered around zero or -0.33, which suggests that state labels were switched (see supplementary material).

@@

A2017.kappa[3, -c(1, seq(2, 13, 3))] <- 0
A2017.kappa[4, seq(2, 13, 3)] <- 0

@@

Comparing gazeHMM to other algorithms on benchmark data showed that gazeHMM showed good agreement with human coders. However, the evaluation criteria (RMSD of event durations, sample-to-sample agreement, and overall disagreement) yielded different results. The fact that gazeHMM outperformed all other algorithms regarding the overall disagreement can be because it is the only algorithm classifying five events (incl. blinks). Nevertheless, Cohen's kappa values of 0.67 (fixations - image) or 0.62 (saccades - moving dots) indicate substantial agreement to human coders, especially in light of the maximum references from the simulation study. At this point it is important to mention that human coding should not be considered a gold standard in event classification: @Hooge2018 observed substantial differences between coders and within coders over time. Despite these differences, they recommend comparisons to human coding to demonstrate the performance of new algorithms and to find errors in their design.

>>>>

library(papaja)
library(tidyverse)
library(signal)
library(CircStats)
library(depmixS4)
library(psych)
library(here)
library(grid)
library(gridExtra)
source(here("algorithm/preprocessing_helper_functions.R"))
source(here("algorithm/model_helper_functions.R"))

@@

On one hand, many of these classification algorithms use non-parametric methods to differentiate between eye movement events. For instance, the classic "Velocity-threshold" algorithm classifies samples with a velocity above a fixed threshold as saccades [@Stampe1993]. On the other hand, many parametric methods have been developed recently. Some of them require human labeled training data as input and can therefore be termed as supervised. For example, @Bellet2019 trained a convolutional neural network (CNN) on eye-tracking data from humans and macaques and achieved classification results that were similarly accurate to those of human coders. However, these supervised methods have several disadvantages: First, they require a sufficient amount of labeled data (usually labeled by humans) as input, which can become costly and time consuming [@Zemblys2019]. Second, they often rely on hyperparameters which are difficult to theoretically justify. Third, most supervised methods operate in a black-box fashion, which implies that they do not make any explicit assumptions about the data generating process [cf. @Zembyls2019]. Beyond that, there are several algorithms using unsupervised techniques. They typically do not need to be trained on labeled data and are also more flexible in classifying data from different individuals, tasks, or eye-trackers [e.g., @Houpt2018; @Hessels2017].  
Besides discriminating between supervised and unsupervised methods, algorithms can vary in whether they are explicitly modeling the data generating process and are thus able to simulate new data. To our knowledge, these *generative models* have been rarely used to classify eye movement data [cf. @Mihali2017; @Wadehn2019]. However, they come with various advantages and we will summarize the two most important ones for eye movement classification here [for an overview, see @Epstein2008].  
First, generative models allow to not only predict (classify) but also explain eye-tracking phenomena. For instance, they can inhere parameters that explain why oscillations only occur after but not before saccades. When the goal is to understand eye movement events and improve their classification based on this understanding, this aspect is clearly an advantage over non-parametric or supervised methods. Second, generative models can challenge prevailing theoretical assumptions and bring up new research questions [@Epstein2008]. For example, they might suggest that oscillations also occur before saccadic eye movements [as has been mentioned by @Nystrom2010] or that the assumption of discrete eye movements does not hold [as has been discussed in @Andersson2017].  

@@

Several studies have tested the performance of HMM algorithms against other classification methods: I-HMM has been deemed as robust against noise, behaviorally accurate, and showing a high sample-to-sample agreement to human coders [@Andersson2017; @Komogortsev2010; @Salvucci2000]. However, the agreement was lower when compared to an algorithm using a Bayesian mixture model [@Kasneci2014; @Tafaj2012]. NSLR-HMM showed even higher agreement to human coding than I-HMM but was outperformed by the CNN algorithm in @Bellet2019 [@Pekkanen2017].  

@@

[^1]: @Hessels2017 called step two the *search rule* and step three the *classifcation rule*. For non-parametric methods, this distinction might be accurate. However, for parametric methods, we think calling step two "classification" is more appropriate since the probabilistic classification is done here. Step three usually consist of some heuristic relabeling and correcting for classification errors.

(ref:plot-algorithm-workflow) Example work flow for eye movement event classification algorithms: (a) the raw gaze signal in x (dark red) and y (dark blue) coordinates; (b) the raw gaze signal is filtered and transformed into a velocity signal; (c) samples are classified as events (indicated by colors), and (d) relabeled. Sequences of samples belonging to the same event are merged (indicated by black segments). Data from @Andersson2017.

```{r plot-algorithm-workflow, fig.cap="(ref:plot-algorithm-workflow)", fig.height = 8}

toPlot <- A2017[[3]][[3]] %>%
  mutate(x.va = px2va(x, dim[1], res[1], dist),
         y.va = px2va(y, dim[2], res[2], dist),
         x.vel = sgolayfilt(x.va, m = 1),
         y.vel = sgolayfilt(y.va, m = 1),
         vel = sqrt(x.vel^2 + y.vel ^2)*fr) %>%
  dplyr::filter(t > 4.3 & t < 4.5) %>%
  mutate(error = vel > 80,
         event.pre = as.factor(ifelse(error, 2, coderMN)),
         event.post = as.factor(coderMN))

# Create plot

p1 <- ggplot(toPlot, aes(x = t, x.va)) + 
  geom_line(aes(y = x.va)) + 
  geom_line(aes(y = y.va)) + 
  geom_point(aes(y = x.va), color = "darkred", size = 0.9) + 
  geom_point(aes(y = y.va), color = "darkblue", size = 0.9) +
  scale_y_continuous(name = "Gaze position (deg)", limits = c(-5, 10), breaks = pretty(c(-5, 10)))

p2 <- ggplot(toPlot, aes(x = t, y = vel)) + 
  geom_line() + 
  geom_point(size = 0.9) +
  scale_y_continuous(name = "Velocity (deg/s)", limits = c(0, 250), breaks = pretty(c(0, 250)))

p3 <- ggplot(toPlot, aes(x = t, y = vel)) + 
  geom_line() + 
  geom_point(aes(color = event.pre), size = 0.9) +
  scale_y_continuous(name = "Velocity (deg/s)", limits = c(0, 250), breaks = pretty(c(0, 250)))

p4 <- ggplot(toPlot, aes(x = t, y = vel)) + 
  geom_line() + 
  geom_point(aes(color = event.post), size = 0.9) +
  scale_y_continuous(name = "Velocity (deg/s)", limits = c(0, 250), breaks = pretty(c(0, 250)))

plotList <- list(p1, p2, p3, p4)

for(p in 1:length(plotList)) {

  plotList[[p]] <- plotList[[p]] + 
    xlab("") + geom_line(color = "black") + ggtitle(letters[p]) + 
    theme(legend.position = "none")

plotList[[4]] <- plotList[[4]] + xlab("Time (s)") +
  geom_segment(aes(x = first(t[event.post == 4]), y = 250, xend = first(t[event.post == 2])-(1/fr), yend = 250), size = 1.2) +
  geom_segment(aes(x = first(t[event.post == 2]), y = 250, xend = first(t[event.post == 3])-(1/fr), yend = 250), size = 1.2) +
  geom_segment(aes(x = first(t[event.post == 3]), y = 250, xend = last(t[event.post == 3]), yend = 250), size = 1.2) +
  geom_segment(aes(x = last(t[event.post == 3])+(1/fr), y = 250, xend = last(t[event.post == 4]), yend = 250), size = 1.2)

grid.arrange(plotList[[1]], plotList[[2]], plotList[[3]], plotList[[4]], ncol = 1)

## Preprocessing
Algorithms require variables that describe gaze data (hereafter called *eye movement metrics*) to classify them into events. Many eye movement metrics have been proposed and used in previous algorithms [for examples, see @Zemblys2018; @Andersson2017]. However, many of them rely on thresholds or window ranges that have to be set by the user [e.g., the distance between the mean position in a 100 ms window before and after each sample, see @Olsson2007]. This can be problematic because such parameters are often set without theoretical justification and differ substantially between metrics or heavily depend on the eye-tracker's sampling frequency [@Andersson2017]. In gazeHMM, we used velocity, acceleration, and sample-to-sample angle [synonymous to relative or change in angle; @Larsson2013] because they belong to the most basic metrics which do not require parameter settings.  
Theoretically, these three metrics should separate eye movement events clearly. Fixations typically inherit samples with low velocity and acceleration [@Larsson2013]. Due to tremor, we assume that the angle between samples should not follow any direction but a uniformly random walk [@Larsson2015]. In contrast, saccade samples usually have a high velocity and acceleration and roughly follow the same direction. PSO samples tend to have moderate velocity and high acceleration since they occur between saccades and low velocity events [@Larsson2013]. They can be specifically distinguished by their change in direction clustered around 180 degrees [@Pekkanen2017]. Importantly, the oscillations depend on the resolution of the gaze recording: Eye-trackers with higher sampling frequency yield more changes in direction and more samples in between those changes. Those samples in between typically follow the same direction. Thus, with high sampling frequencies, PSO samples might also cluster around a sample-to-sample angle of zero with outliers around 180 degrees. Lastly, smooth pursuit samples have a moderate velocity but low acceleration (due to the smoothness) and like saccades they follow the same direction [@Larsson2013; @Leigh2015]. Other algorithms focus exclusively on classifying microsaccades [e.g., @Mihali2017], but as stated earlier, these events were not in the scope of gazeHMM.  
The velocity and acceleration signals are computed from the raw gaze position by using a Savitzky-Golay filter [@Savitzky1964; similar to @Nystrom2010]. The sample-to-sample angle is calculated as: $$\alpha(t) = \arctan\left(\frac{y_{t+1} - y_{t}}{x_{t+1} - x_{t}}\right) -\arctan\left(\frac{y_{t} - y_{t-1}}{x_{t} - x_{t-1}}\right),$$ with $\alpha(t) := \alpha(t) + 2\pi$ for $\alpha(t) < 0$, and is therefore bound between 0 and $2\pi$.  
Most of the missing data in eye movement classification are due to blinks. Therefore, the user can indicate which samples should be labeled as blinks (other missing samples are treated as noise). Often, eye-trackers record a few samples with unreasonably high velocity and acceleration before losing the pupil signal when a blink occurs. Since these samples could distort the classification of saccades in the HMM, gazeHMM removes them heuristically. Before classifying the samples, it sets all samples within 50 ms before and after blink samples as missing. The window of 50 ms was rather motivated empirically than theoretically and can bet set to any value the user considers appropriate.

@@

When we developed gazeHMM, the gamma distribution appeared to fit eye movement data well, but we also note that it might not necessarily be the best fitting distribution for every type of eye movement data. We assume that the best fitting distribution will depend on the task, eye-tracker, and individual (see discussion).  

@@

Both the distributions and the metric operate on the full unit circle (i.e., between 0 and $2\pi$), which should lead to rather symmetric distributions. Because we assume fixations to change their direction similar to a uniformly random walk [@Larsson2013; @Larsson2015], their sample-to-sample angle can be modeled by a uniform distribution. Thus, the uniform distribution should distinguish fixations from the other events.  

@@

## Postprocessing
After classifying gaze samples into states, gazeHMM applies a postprocessing routine to the estimated state sequence. We implemented this routine because constraining the probabilities for nonsaccade events to turn into PSOs to zero often caused PSOs not to appear in the state sequence at all. Moreover, gazeHMM does not explicitly control the duration of events in the HMM which occasionally led to unreasonably short events. Thus, the postprocessing routine heuristically compensates for such violations. This routine relabels one-sample fixations and smooth pursuits, saccades with a duration below a minimum threshold, and PSOs that follow nonsaccade events. Samples are relabeled as the state of the previous event. Finally, samples initially indicated as missing are labeled as noise (including blinks) and event descriptives are computed (e.g., fixation duration).  
The algorithm is implemented in R [version: 3.6.3; @RCoreTeam2020] and uses the packages \texttt{signal} [@Ligges2015] to compute velocity and acceleration signals, \texttt{depmixS4} [@Visser2010] for the HMM, \texttt{CircStats} [@Lund2018] for the von Mises distribution, and \texttt{BB} [@Varadhan2009] for Barzilai-Borwein spectral projected gradient optimization. The algorithm is available on GitHub (www.github.com/maltelueken/gazeHMM).

@@

## Results
### Parameter Variation
In the first part of the simulation, we examined how varying the parameters[^2] in the HMM affected the deviation of estimated parameters and accuracy of estimated state sequences. For the two-state HMM, the recovery of parameters and states was nearly perfect (all RMdSPDs < 0.1, intercepts and slopes of regression lines almost zero and one, respectively, and Cohen's kappa close to 1). We therefore chose to include the respective figures in the supplementary material.  
For the HMM with three states, the RMdSPD is shown in Figure \@ref(fig:plot-rmdspd-1-3). When response parameters (other than $a_{i=j}$) were manipulated, the RMdSPDs for $a_{12}$ and $a_{31}$ were consistently between 0.1 and 0.5. Varying $\kappa$ in states two and three led to RMdSPDs between 0.1 and 0.5 in the respective states, which we interpreted as moderate recovery. Otherwise, RMdSPDs were consistently lower than 0.1, indicating good recovery. Inspecting the regression lines between true and estimated parameters (see Figures \@ref(fig:plot-regtr-1-3) and \@ref(fig:plot-regresp-1-3)) revealed strong and unbiased linear relationships (intercepts close to zero and slopes close to one). In contrast to the two-state HMM, larger deviations and more outliers were observed. Cohen's kappa values are presented in Figure \@ref(fig:plot-acc-1-3). For most estimated models, the kappa values between true and estimated state sequences were above 0.95, meaning almost perfect agreement. However, for some models, we observed kappas clustered around zero or -0.33, which suggests that state labels were switched (see supplementary material).  
[^2]: Note that the initial state probability $\rho_i$ has $\textrm{RMdSPD} = 1$. Since the HMM only simulated one state sequence, this parameter is always either zero or one (leading to $\textrm{RMdSPD} = 1$). Therefore, we decided not to include it in the analysis.

@@

A2017.kappa[3, -c(1, seq(2, 13, 3))] <- NA
A2017.kappa[4, seq(2, 13, 3)] <- NA

@@

Comparing gazeHMM to other algorithms on benchmark data showed that gazeHMM showed good agreement with human coders. However, the evaluation criteria (RMSD of event durations, sample-to-sample agreement, and overall disagreement) yielded different results. The fact that gazeHMM outperformed all other algorithms regarding the overall disagreement can be because it is the only algorithm classifying five events [incl. blinks; see also the *accuracy paradox*, @Zemblys2019]. Nevertheless, Cohen's kappa values of 0.67 (fixations - image) or 0.62 (saccades - moving dots) indicate substantial agreement to human coders, especially in light of the maximum references from the simulation study. At this point it is important to mention that human coding should not be considered a gold standard in event classification: @Hooge2018 observed substantial differences between coders and within coders over time. Despite these differences, they recommend comparisons to human coding to demonstrate the performance of new algorithms and to find errors in their design.
